Convex Optimization: A Very Short and Incomplete Summary

### 2013-10-02 icml/convex-opt-summary.tex

The algorithms, along with conditions on the functions and convergence rates, are summarized in the table on the next page. Here we define some of the notation used in specific algorithms. In projected gradient descent $\eta_{t} = 2/\alpha(t+1)$. In Nesterov's accelerated gradient descent,
\[
\lambda_{0} &= 0 \\
\lambda_{t} &= \frac{1 + \sqrt{1 + 4\lambda_{t-1}^{2}}}{2} \\
\gamma_{t}  &= \frac{1 - \lambda_{t}}{\lambda_{t+1}}.
\]
In mirror descent and conditional gradient descent $\rho_{t} = 2/(t + 1)$. Finally, in boosted mirror descent
\[
\alpha_{s} &= s \\
A_{t} &= \sum_{s=1}^{t} \alpha_{s} = t(t+1)/2 \\
\bar x_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}x_{s} \\
\bar y_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}y_{s}
\]

@@

\begin{table}[htdp]
\begin{center}
\begin{tabular}{|l|p{4.05cm}|p{4.05cm}|p{4.05cm}|p{4.05cm}|}
\hline & Gradient Descent & Projected Subgrad. Desc. & Conditional Grad. Desc. & Nesterov's Accelerated Grad. Desc.   \\ 
\hline
Objective function  &
$f(x)$ &
$f(x)$ &
$-h^{*}(y) - f^{*}(y)$ &
$f(x)$ \\
\hline
Conditions & 
$f$ is $\beta$-smooth &
$f$ is $\alpha$-s.c., $\|\partial f\| \le L$, and space is compact. & 
$f$ is L.c. and $h$ is $\alpha$-s.c. &
$f$ is $\beta$-smooth \\
\hline
Update equations &
$x_{t+1} = x_{t} - \frac{1}{\beta}\partial f(x_{t})$ &
$y_{t+1} = x_{t} - \eta_{t}\partial f(x_{t})$ and $x_{t+1} = \argmax_{x} \| x - y_{t+1} \|$ &
$x_{t} = \partial h^{*}(-y_{t})$ and $\bar y_{t} = \partial f(x_{t})$ and $y_{t+1} = (1 - \rho_{t})y_{t} + \rho_{t} \bar y_{t}$&
$y_{t+1} = x_{t} - \frac{1}{\beta}\partial f(x_{t})$ and $x_{t+1} = (1- \gamma_{t})y_{t+1} + \gamma_{t}y_{t}$ \\
\hline
Convergence rate &
$\beta R^{2}/t$ &
$L^{2} / \alpha t$ &
$R^{2} / \alpha t$ &
$\beta R^{2}/t^{2}$ \\
\hline
\end{tabular}
\caption{Convex optimization algorithms (gradient descent type). The function $f$ and $g$ are assumed to be convex. Constants are not given in convergence rates. Here $R$ is the radius of the metric space. Shorthands: L.c. = Lipschitz continuous, s.c. = strongly convex. }
\label{table:gd-convex-opt-algs}
\end{center}
\end{table}

\begin{table}[htdp]
\begin{center}
\begin{tabular}{|l|p{5.5cm}|p{5.5cm}|p{5.5cm}|}
\hline & Mirror Descent & ``Primal'' Boosted Mirror Descent & ``Dual'' Boosted Mirror Descent  \\ 
\hline
Objective function & 
$h(x) + f(x)$ & 
$h(x) + f^{*}(x)$ & 
$h(x) + f^{*}(x)$ \\
\hline
Conditions & 
$h$ is smooth \& $\alpha$-s.c. and $f$ is L.c. & 
$h$ is $\alpha$-s.c. and $f^{*}$ is convex& 
$f^{*}$ is $\beta$-smooth and covex \\
\hline
Update equations & 
$y_{t} = \partial f(x_{t})$ and 
$x_{t+1} = \argmin_{x} h(x) - (1 - \rho_{t})\ip{\partial h(x_{t})}{x} + \rho_{t+1}\ip{y_{t}}{x}$ &
$x_{t} = \partial h^{*}(-\bar y_{t})$ and $y_{t+1} = \partial f^{*}(x_{t})$ & 
$x_{t} = \partial h^{*}(-y_{t})$ and $y_{t+1} = \partial f^{*}(\bar x_{t})$ \\
\hline
Convergence rate & 
$R^{2}/\alpha t$ & 
$R^{2}/\alpha t$ & 
$\beta R^{2} / t$ \\
\hline

@@

\end{landscape}
\end{document}

>>>>

The algorithms, along with conditions on the functions and convergence rates, are summarized in the table on the next page. Here we define some of the notation used in specific algorithms. In projected gradient descent $\eta_{t} = 2/\alpha(t+1)$. In Nesterov's accelerated gradient descent,
\[
\lambda_{0} &= 0 \\
\lambda_{t} &= \frac{1 + \sqrt{1 + 4\lambda_{t-1}^{2}}}{2} \\
\gamma_{t}  &= \frac{1 - \lambda_{t}}{\lambda_{t+1}}.
\]
In mirror descent and conditional gradient descent $\rho_{t} = 2/(t + 1)$. Finally, in boosted mirror descent
\[
\alpha_{s} &= s \\
A_{t} &= \sum_{s=1}^{t} \alpha_{s} = t(t+1)/2 \\
\bar u_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}u_{s} \\
\bar \theta_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}\theta_{s}
\]

@@

\begin{table}[htdp]
\begin{center}
\begin{tabular}{|l|p{4.05cm}|p{4.05cm}|p{4.05cm}|p{4.05cm}|}
\hline & Gradient Descent & Projected Subgrad. Desc. & Conditional Grad. Desc. & Nesterov's Accelerated Grad. Desc.   \\ 
\hline
Objective function  &
$f(x)$ &
$f(x)$ &
$h^{*}(y) + f^{*}(y)$ &
$f(x)$ \\
\hline
Conditions & 
$f$ is $\beta$-smooth &
$f$ is $\alpha$-s.c., $\|\partial f\| \le L$, and space is compact. & 
$f$ is L.c. and $h$ is $\alpha$-s.c. &
$f$ is $\beta$-smooth \\
\hline
Update equations &
$x_{t+1} = x_{t} - \frac{1}{\beta}\partial f(x_{t})$ &
$y_{t+1} = x_{t} - \eta_{t}\partial f(x_{t})$ and $x_{t+1} = \argmax_{x} \| x - y_{t+1} \|$ &
$x_{t} = \partial h^{*}(-\bar y_{t})$ and $y_{t+1} = \partial f(x_{t})$ and $\bar y_{t+1} = (1 - \rho_{t})\bar y_{t} + \rho_{t} y_{t+1}$&
$y_{t+1} = x_{t} - \frac{1}{\beta}\partial f(x_{t})$ and $x_{t+1} = (1- \gamma_{t})y_{t+1} + \gamma_{t}y_{t}$ \\
\hline
Convergence rate &
$\beta R^{2}/t$ &
$L^{2} / \alpha t$ &
$R^{2} / \alpha t$ &
$\beta R^{2}/t^{2}$ \\
\hline
\end{tabular}
\caption{Convex optimization algorithms (gradient descent type). The function $f$ and $g$ are assumed to be convex. Constants are not given in convergence rates. Here $R$ is the radius of the metric space. Shorthands: L.c. = Lipschitz continuous, s.c. = strongly convex. }
\label{table:gd-convex-opt-algs}
\end{center}
\end{table}

\begin{table}[htdp]
\begin{center}
\begin{tabular}{|l|p{5.5cm}|p{5.5cm}|p{5.5cm}|}
\hline & Mirror Descent & ``Primal'' Boosted Mirror Descent & ``Dual'' Boosted Mirror Descent  \\ 
\hline
Objective function & 
$h(x) + f(x)$ & 
$g(x) + R^{*}(x)$ & 
$g(x) + R^{*}(x)$ \\
\hline
Conditions & 
$h$ is smooth \& $\alpha$-s.c. and $f$ is L.c. & 
$g$ is $\alpha$-s.c.& 
$R^{*}$ is $\beta$-smooth \\
\hline
Update equations & 
$y_{t} = \partial f(x_{t})$ and 
$x_{t+1} = \argmin_{x} h(x) - (1 - \rho_{t+1})\ip{\partial h(x_{t})}{x} + \rho_{t+1}\ip{y_{t}}{x}$ &
$u_{t} = \partial g^{*}(-\bar \theta_{t})$ and $\theta_{t+1} = \partial R^{*}(u_{t})$ & 
$u_{t} = \partial g^{*}(-\theta_{t})$ and $\theta_{t+1} = \partial R^{*}(\bar u_{t})$ \\
\hline
Convergence rate & 
$R^{2}/\alpha t$ & 
$R^{2}/\alpha t$ & 
$\beta R^{2} / t$ \\
\hline

@@

\end{landscape}

\section{CGD, MD, and Boosted MD}

The Boosted Mirror Descent algorithms are very closely related to Conditional Gradient Descent and Mirror Descent. Dual BMD and CGD are related as follows. BMD minimizes $g + R^{*}$ while CGD minimizes $h^{*} + f^{*}$. Make the identifications $h^{*} = R^{*}$ and $f^{*} = g$. Note that $R^{*}$ is $\beta$-smooth $\iff R$ is $(1/\beta)$-strongly convex $\iff h$ is $(1/\beta)$-strongly convex, so the conditions on the functions being optimized are equivalent. Then with $\rho_{t} = \alpha_{t+1}/A_{t+1}$, we can rewrite the CGD updates  as 
\[
x_{t} &= \partial R^{*}(-\bar y_{t})  \\
y_{t} &= \partial g^{*}(x_{t}) \\
\bar y_{t+1} &= A_{t+1}^{-1}\sum_{s \le t+1} \alpha_{t} y_{t},
\]
which is identical to the dual BMD if we make the substitutions $x_{t} \to -\theta_{t}$ and $y_{t} \to -u_{t}$.

We similarly relate Primal BMD and Mirror Descent. BMD minimizes $g + R^{*}$ while CGD minimizes $h + f$. Make the identifications $h = g$ and $f = R^{*}$ and note the conditions of $h$ and $g$ are the same. Letting $\rho_{t+1} = \alpha_{t}/A_{t}$, we can rewrite the Mirror Descent updates as 
\[
y_{t} &= \partial R^{*}(x_{t}) \\
x_{t+1} &= \partial g^{*}((1 - \rho_{t+1})\partial g(x_{t}) - \rho_{t+1}y_{t}) \\
&= \partial g^{*}(-\bar y_{t}), 
\]
where $\bar y_{t}$ is the same as above. Making the substitutions $y_{t} \to \theta_{t}$ and $x_{t} \to u_{t}$ completes the identification of the two methods.

\end{document}
